---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(leaps)
library(car)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
load("movies.Rdata")
```

* * *

## Part 1: Data
- The data is a random sample collected from IMDB and Rotten Tomatoes on various
movies and attributes associated with them.
- The scope of inference is limited to generalization as this is an observational
study, not a randomized controlled trial. No causal connections can be made from
this analysis.

* * *

## Part 2: Research question
- What attributes of a movie are most strongly associated with audience score?

* * *

## Part 3: Exploratory data analysis

#### Variables being considered
Dependent Variable: 
- audience_score

Independent Variables: 
- critics_score
- critics_rating
- title_type
- genre
- runtime
- mpaa_rating
- best_pic_nom
- best_pic_win
- best_actor_win
- best_actress_win
- best_dir_win
- top200_box

### Investigate the Dataset and Variables
```{r}
# Investigate the Dataset
head(movies)
tail(movies)
str(movies)
summary(movies)

## Investigate the variables

# Dependent Variable
summary(movies$audience_score)
hist(movies$audience_score)
boxplot(movies$audience_score)
scatter.smooth(movies$audience_score)
qqnorm(movies$audience_score)
qqline(movies$audience_score)

# Independent Variables

# critics_rating
summary(movies$critics_rating)
barplot(table(movies$critics_rating))

# audience_rating
summary(movies$audience_rating)
barplot(table(movies$audience_rating))

# critics_score
summary(movies$critics_score)
hist(movies$critics_score)
boxplot(movies$critics_score)
scatter.smooth(movies$critics_score)
qqnorm(movies$critics_score)
qqline(movies$critics_score)

# imdb_rating
summary(movies$imdb_rating)
hist(movies$imdb_rating)
boxplot(movies$imdb_rating)
scatter.smooth(movies$imdb_rating)
qqnorm(movies$imdb_rating)
qqline(movies$imdb_rating)

# imdb_num_votes
summary(movies$imdb_num_votes)
hist(movies$imdb_num_votes)
boxplot(movies$imdb_num_votes)
scatter.smooth(movies$imdb_num_votes)
qqnorm(movies$imdb_num_votes)
qqline(movies$imdb_num_votes)

# genre
summary(movies$genre)
barplot(table(movies$genre))

# runtime
summary(movies$runtime)
hist(movies$runtime)
boxplot(movies$runtime)
scatter.smooth(movies$runtime)
qqnorm(movies$runtime)
qqline(movies$runtime)

# mpaa_rating
summary(movies$mpaa_rating)
barplot(table(movies$mpaa_rating))

# best_pic_nom
summary(movies$best_pic_nom)
barplot(table(movies$best_pic_nom))

# best_pic_win
summary(movies$best_pic_win)
barplot(table(movies$best_pic_win))

# best_actor_win
summary(movies$best_actor_win)
barplot(table(movies$best_actor_win))

# best_actress_win
summary(movies$best_actress_win)
barplot(table(movies$best_actress_win))

# best_dir_win
summary(movies$best_dir_win)
barplot(table(movies$best_dir_win))

# top200_box
summary(movies$top200_box)
barplot(table(movies$top200_box))

## Determine which Predictor (Independent) variables are most highly correlated
```
### Data Manipulation and Cleaning
```{r}
# Keep columns for analysis
movies_trimmed <- movies[, c("audience_score","genre", "runtime","mpaa_rating","title_type","critics_score","best_pic_nom","best_pic_win","best_actor_win","best_actress_win","best_dir_win","top200_box","imdb_rating")]

# Check rows with N/A's and remove if feasible
num_rows_with_na <- sum(is.na(movies_trimmed))

# Remove rows with N/As.
movies_cleaned <- na.omit(movies_trimmed)

# Analyze cleaned data
str(movies_cleaned)
```

* * *

## Part 4: Modeling

### Backwards Elimination - P-Value

```{r}
## Multiple Linear Regression Model

# Full model
score_full = lm(audience_score ~ critics_score + title_type + genre
                + runtime + mpaa_rating + best_pic_nom + best_pic_win + best_actor_win + 
                  best_actress_win + best_dir_win + top200_box + imdb_rating, data = movies_cleaned)
summary(score_full)

# Model - mpaa_rating
score_mod_1 = lm(audience_score ~ critics_score + title_type + genre
                + runtime + best_pic_nom + best_pic_win + best_actor_win + 
                  best_actress_win + best_dir_win + top200_box + imdb_rating, data = movies_cleaned)
summary(score_mod_1)

# Model - best_pic_win
score_mod_2 = lm(audience_score ~ critics_score + title_type + genre
                + runtime + best_pic_nom + best_actor_win + 
                  best_actress_win + best_dir_win + top200_box + imdb_rating, data = movies_cleaned)
summary(score_mod_2)

# Model - title_type
score_mod_3 = lm(audience_score ~ critics_score + genre
                + runtime + best_pic_nom + best_actor_win + 
                  best_actress_win + best_dir_win + top200_box + imdb_rating, data = movies_cleaned)
summary(score_mod_3)

# Model - best_dir_win
score_mod_4 = lm(audience_score ~ critics_score + genre
                + runtime + best_pic_nom + best_actor_win + 
                  best_actress_win + top200_box + imdb_rating, data = movies_cleaned)
summary(score_mod_4)

# Model - top200_box
score_mod_5 = lm(audience_score ~ critics_score + genre
                + runtime + best_pic_nom + best_actor_win + 
                  best_actress_win + imdb_rating, data = movies_cleaned)
summary(score_mod_5)

# Model - best_actor_win
score_mod_6 = lm(audience_score ~ critics_score + genre
                + runtime + best_pic_nom + 
                  best_actress_win + imdb_rating, data = movies_cleaned)
summary(score_mod_6)

# Model - best_actress_win
score_mod_7 = lm(audience_score ~ critics_score + genre
                + runtime + best_pic_nom + 
                  imdb_rating, data = movies_cleaned)
summary(score_mod_7)


# Model - best_pic_nom
score_mod_8 = lm(audience_score ~ critics_score + genre
                + runtime +  
                  imdb_rating, data = movies_cleaned)
summary(score_mod_8)

# Final Model
score_full = lm(audience_score ~ critics_score + genre
                + runtime +  
                  imdb_rating, data = movies_cleaned)
summary(score_full)
```

### Assess Collinearity of Predictor Variables in the Final Model
The adjusted GVIF taking into account the DF indicates a relatively low indication
of collinearity between predictor (independent) variables.
However, the unadjusted GVIF is approaching the threshold of acceptable level of
collinearity (5-10). 

```{r}
## Correlation matrix could not be used due to the presence of categorical variables

# Calculate VIF for the predictors in the model
vif_values <- vif(score_full)

# Print the VIF values
print(vif_values)
```

### Conditions

#### (1) linear relationships between (numerical) x and y
Check if each explanatory variable is linearly related to the response variable.

‣ Each (numerical) explanatory variable linearly related to the response variable
‣ Check using residuals plots (e vs. x)
‣ Looking for a random scatter around 0
‣ Instead of scatterplot of y vs. x: allows for considering the other variables 
that are also in the model, and not just the bivariate relationship between a 
given x and y.

```{r}
plot(score_full$residuals ~ movies_cleaned$critics_rating)
plot(score_full$residuals ~ movies_cleaned$critics_score)
plot(score_full$residuals ~ movies_cleaned$title_type)
plot(score_full$residuals ~ movies_cleaned$genre)
plot(score_full$residuals ~ movies_cleaned$runtime)
plot(score_full$residuals ~ movies_cleaned$mpaa_rating)
plot(score_full$residuals ~ movies_cleaned$best_pic_nom)
plot(score_full$residuals ~ movies_cleaned$best_pic_win)
plot(score_full$residuals ~ movies_cleaned$best_actor_win)
plot(score_full$residuals ~ movies_cleaned$best_actress_win)
plot(score_full$residuals ~ movies_cleaned$best_dir_win)
plot(score_full$residuals ~ movies_cleaned$top200_box)
plot(score_full$residuals ~ movies_cleaned$imdb_rating)
```

#### (2) nearly normal residuals with mean 0
The plots below indicate a nearly normal distribution of residuals scattered
around zero within the model.

‣ Some residuals will be positive and some negative
‣ On a residuals plot we look for random scatter of residuals around 0
‣ This translates to a nearly normal distribution of residuals centered at 0
‣ Check using histogram or normal probability plot

```{r}
hist(score_full$residuals)
qqnorm(score_full$residuals)
qqline(score_full$residuals)
```
#### (3) constant variability of residuals
Constant variance is effectively met, with a slight convergence indicated nearing
the right side of the residuals plot.

‣ Residuals should be equally variable for low and high values of the predicted 
response variable
‣ Check using residuals plots of residuals vs. predicted (e vs. )
‣ Residuals vs. predicted instead of residuals vs. x because it allows for 
considering the entire model (with all explanatory variables) at once
‣ Residuals randomly scattered in a band with a constant width around 0 (no 
fan shape)
‣ Also worthwhile to view absolute value of residuals vs. predicted to identify 
unusual observations easily

```{r}
plot(score_full$residuals ~ score_full$fitted)
plot(abs(score_full$residuals) ~ score_full$fitted)
```
#### (4) independent residuals
It does not appear there is time series structure involved with this data. The
data is a random sample of movies from IMBD and Rotten Tomatoes so independence should
be satisfied.

‣ Independent residuals → independent observations
‣ If time series structure is suspected check using residuals vs. order of data 
collection
‣ If not, think about how the data are sampled

```{r}
plot(score_full$residuals)
```

* * *

## Part 5: Prediction

```{r}
# Calculate confidence intervals for the coefficients
conf_intervals <- confint(score_full)

# View the confidence intervals
conf_intervals

## Determine if the movie being used in the model exists within the dataset

# Find rows where the string "Rogue" exists in the 'Name' column
matching_rows <- movies[grepl("", movies$title, ignore.case = TRUE), ]

# Print the matching rows
matching_rows

# Final Model
score_full = lm(audience_score ~ critics_score + genre
                + runtime +  
                  imdb_rating, data = movies_cleaned)
summary(score_full)

## Acquire data from various movies released in 2016

# Data from Rogue One
rogue_one = data.frame(critics_score = c(84), genre = c("Action & Adventure"), runtime = c(134), imdb_rating = c(7.8))

# Data from Batman vs Superman
batman_vs_superman = data.frame(critics_score = c(29), genre = c("Action & Adventure"), runtime = c(153), imdb_rating = c(6.4))

# Data from X-Men: Apocalypse
xmen_apocalypse = data.frame(critics_score = c(47), genre = c("Action & Adventure"), runtime = c(143), imdb_rating = c(6.9))

## Make predictions based on the movie data above
rogue_one_predictions <- predict(score_full, rogue_one)
batman_vs_superman_predictions <- predict(score_full, batman_vs_superman)
xmen_apocalypse_predictions <- predict(score_full, xmen_apocalypse)

## View the predicted values
rogue_one_predictions 
batman_vs_superman_predictions
xmen_apocalypse_predictions
```


* * *

## Part 6: Conclusion
Utilizing data acquired from other movies released in 2016 not in the dataset,
the model appears to do a decent job of predicting the audience score. After improving
the model with another iteration, the model can now predict Rotten Tomatoes audience
scores within a smaller confidence interval. However, the improved model may
violate the constant variance requirement. Another improvement to the model can be made,
such as identifying if any other variables can be removed/added.
